---
title: "Appendix A3: literature data treatment."
output:
  pdf_document:
    toc: true
    toc_depth: 3
bibliography: references.bib
---

\setcounter{table}{0}
\renewcommand{\thetable}{C.\arabic{table}}
\renewcommand{\thefigure}{C.\arabic{figure}}


# Fishing gear

## Coastal fishers

Professional fisherman fishing effort is available at detailed level in in the Fishery Dependent Information (FDI) data from the STECF [@stecf2024]. FDI data are filtered on the Swedish Bothnian Sea by filtering on sub region (27.30.D.30) and country code (SWE). Data are further filtered on the vessel length categories < 8m and 8-12m length overall.

Coastal fishers adopt a large variety of fishing technique. We selected the three most important in number of fishing days, which also accounts for more than 85%: 

```{r, include=T, echo=F,message=FALSE, results = 'asis', fig.width = 6, fig.height=3, warning=FALSE, fig.cap='The 8 combination of gear and target species having the most fishing days over the period 2013-2023.'}
library(tidyverse)
x.eff=read_csv("~/ARMELLONI_SLU/miscellaneous_datasets/2024_fdi_effort/Effort/FDI Effort by country.csv")
names(x.eff)=str_replace(names(x.eff), ' ', '_')
names(x.eff)=str_replace(names(x.eff), ' ', '_')
names(x.eff)=str_replace(names(x.eff), '-', '_')
names(x.eff)=tolower(names(x.eff))

eff.30=x.eff[x.eff$sub_region=='27.3.D.30' &x.eff$country=='Sweden',]
coastal.eff=eff.30[eff.30$vessel_length_category %in% c('VL0812', 'VL0008'),]

## Decision
coastal_fdays=coastal.eff%>%
  dplyr::mutate(total_fishing_days=as.numeric(total_fishing_days))%>%
  dplyr::group_by(fishing_technique,gear_type,target_assemblage,year)%>%
  dplyr::summarise(fdays=sum(total_fishing_days))%>%
  dplyr::group_by(fishing_technique,gear_type, target_assemblage)%>%
  dplyr::summarise(fdays=mean(fdays))%>%
  dplyr::group_by(gear_type, target_assemblage)%>%
  dplyr::summarise(fdays=sum(fdays))%>%
  arrange(desc(fdays))%>%
  ungroup()%>%
  dplyr::mutate(cum_val=cumsum(fdays)/sum(fdays), prop=fdays/sum(fdays)) ## CPT for fishing_style -> gear

coastal_fdays$metier=paste(coastal_fdays$gear_type, coastal_fdays$target_assemblage, sep='_')
#write.csv(coastal_fdays, "C:/github/extreme_weather_risk/data/read_only/coastal_fdays.csv", row.names = F)

coastal_fdays[1:8,]%>%
  ggplot(aes(x=reorder(metier, desc(fdays)), y=fdays))+
  geom_col()+
  xlab('Gear & Target assemblage')

```


\newpage

# Target species

Professional fisherman catchyes by gear is available at detailed level in in the Fishery Dependent Information (FDI) data from the STECF [@stecf2024]. FDI data are filtered on the Swedish Bothnian Sea by filtering on sub region (27.30.D.30) and country code (SWE).

Recreational fishers statistics are available in the recreational fisheries statistical report of the Swedish Department for Economic and Statistic Analysis (SWAM) data source [@swam2024]. SWAM target species data are available for coastal vs inland fishing, however they represent the totality of recreational fisherman in Sweden, with no distinction between region.


## Coastal fishers

FDI data are further filtered on the vessel length categories < 8m and 8-12m length overall, and on the three combination of fishing gear and target assemblage selected above (hereby referred as metier).

We calculated for each metier the species assemblage making up to the 95% of the average cumulative distribution in the period 2013-2023: 

```{r, include=T, echo=F, results = 'asis', message=FALSE, fig.width = 8, fig.height=4, warning=FALSE, fig.cap='CPT estimated for pelagic trawlers'}
asfis= read_csv("~/ARMELLONI_SLU/miscellaneous_datasets/ASFIS_2020.csv")
names(asfis)[3]='species'
asfis=asfis[,3:5]
fdi.folder='~/ARMELLONI_SLU/miscellaneous_datasets/2024_fdi_catches/Catches'
fdi.dats=list.files(fdi.folder)
fdi.dats=fdi.dats[nchar(fdi.dats)<40]
fdi.yrs=length(fdi.dats)

fdi.store=NULL
for(i in 1:fdi.yrs){
  fdi=read_csv(file.path(fdi.folder, fdi.dats[i]))
  fdi.30=fdi[fdi$sub_region=='27.3.D.30' &fdi$country=='Sweden' & fdi$vessel_length %in% c('VL0812', 'VL0008'),]
  fdi.store=rbind(fdi.store, fdi.30)
}

fdi.30=fdi.store%>%
  dplyr::mutate(tons=as.numeric(total_live_weight_landed))%>%
  dplyr::filter(!is.na(tons))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, species, year)%>%
  dplyr::summarise(ton=sum(tons))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, species)%>%
  dplyr::summarise(ton=mean(ton))
fdi.30$metier=paste(fdi.30$gear_type, fdi.30$target_assemblage, sep='_')

coastal.catch=fdi.30[fdi.30$metier %in% coastal_fdays[1:3,]$metier ,]%>%
  dplyr::group_by(metier, species)%>%
  dplyr::summarise(ton=sum(ton))%>%
  dplyr::arrange(metier, desc(ton))%>%
  dplyr::mutate(cum_val=round(cumsum(ton)/sum(ton), digits=2), prop=ton/sum(ton))

coastal.catch$keep=1
for(i in 2:(nrow(coastal.catch))){
   if(coastal.catch[i,]$metier==coastal.catch[i-1,]$metier){
     if(coastal.catch[i,]$cum_val<=0.95){
       next
     }else{
       coastal.catch[i,]$keep=0
     }
   }else{
     coastal.catch[i,]$keep=1
   }
}
coastal.catch=coastal.catch%>%
  dplyr::filter(keep==1)%>%
  left_join(asfis)

coastal.catch%>%
  ggplot(aes(x=English_name, y=ton))+
  geom_col()+
  facet_wrap(~metier, scales='free')+
  xlab('Species name')

#write.csv(coastal.catch, "C:/github/extreme_weather_risk/data/read_only/coastal_catch.csv", row.names = F)
```

## Pelagic trawlers

FDI data are further filtered on the vessel length categories 24-40m and >40m length overall, and on the fishing gear and target assemblage combination OTM_SPF.

We calculated the species assemblage making up to the 99% of the average cumulative distribution in the period 2013-2023: 

```{r, include=T, echo=F, results = 'asis', message=FALSE, fig.width = 8, fig.height=4, warning=FALSE, fig.cap='CPT estimated for pelagic trawlers'}
asfis= read_csv("~/ARMELLONI_SLU/miscellaneous_datasets/ASFIS_2020.csv")
names(asfis)[3]='species'
asfis=asfis[,3:5]
fdi.folder='~/ARMELLONI_SLU/miscellaneous_datasets/2024_fdi_catches/Catches'
fdi.dats=list.files(fdi.folder)
fdi.dats=fdi.dats[nchar(fdi.dats)<40]
fdi.yrs=length(fdi.dats)

fdi.store=NULL
for(i in 1:fdi.yrs){
  fdi=read_csv(file.path(fdi.folder, fdi.dats[i]))
  fdi.30=fdi[fdi$sub_region=='27.3.D.30' &fdi$country=='Sweden' & fdi$vessel_length %in% c('VL2440', 'VL40XX'),]
  fdi.store=rbind(fdi.store, fdi.30)
}

fdi.30=fdi.store%>%
  dplyr::mutate(tons=as.numeric(total_live_weight_landed))%>%
  dplyr::filter(!is.na(tons))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, species, year)%>%
  dplyr::summarise(ton=sum(tons))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, species)%>%
  dplyr::summarise(ton=mean(ton))
fdi.30$metier=paste(fdi.30$gear_type, fdi.30$target_assemblage, sep='_')

coastal.catch=fdi.30[fdi.30$metier =='OTM_SPF' ,]%>%
  dplyr::group_by(metier, species)%>%
  dplyr::summarise(ton=sum(ton))%>%
  dplyr::arrange(metier, desc(ton))%>%
  dplyr::mutate(cum_val=round(cumsum(ton)/sum(ton), digits=2), prop=ton/sum(ton))

coastal.catch$keep=1
for(i in 2:(nrow(coastal.catch))){
   if(coastal.catch[i,]$metier==coastal.catch[i-1,]$metier){
     if(coastal.catch[i,]$cum_val<=0.99){
       next
     }else{
       coastal.catch[i,]$keep=0
     }
   }else{
     coastal.catch[i,]$keep=1
   }
}
coastal.catch=coastal.catch%>%
  dplyr::filter(keep==1)%>%
  left_join(asfis)

coastal.catch%>%
  ggplot(aes(x=English_name, y=ton))+
  geom_col()+
  facet_wrap(~metier, scales='free_x')+
  xlab('Species name')

#write.csv(coastal.catch, "C:/github/extreme_weather_risk/data/read_only/coastal_catch.csv", row.names = F)
```


## Recreational fishers

FDI data are further filtered on the vessel length categories 24-40m and >40m length overall, and on the fishing gear and target assemblage combination OTM_SPF.

We calculated the species assemblage making up to the 99% of the average cumulative distribution in the period 2013-2023: 

```{r, include=T, echo=F, results = 'asis', message=FALSE, fig.width = 8, fig.height=4, warning=FALSE, fig.cap='CPT estimated for pelagic trawlers'}
asfis= read_csv("~/ARMELLONI_SLU/miscellaneous_datasets/ASFIS_2020.csv")
names(asfis)[3]='species'
asfis=asfis[,3:5]
fdi.folder='~/ARMELLONI_SLU/miscellaneous_datasets/2024_fdi_catches/Catches'
fdi.dats=list.files(fdi.folder)
fdi.dats=fdi.dats[nchar(fdi.dats)<40]
fdi.yrs=length(fdi.dats)

fdi.store=NULL
for(i in 1:fdi.yrs){
  fdi=read_csv(file.path(fdi.folder, fdi.dats[i]))
  fdi.30=fdi[fdi$sub_region=='27.3.D.30' &fdi$country=='Sweden' & fdi$vessel_length %in% c('VL2440', 'VL40XX'),]
  fdi.store=rbind(fdi.store, fdi.30)
}

fdi.30=fdi.store%>%
  dplyr::mutate(tons=as.numeric(total_live_weight_landed))%>%
  dplyr::filter(!is.na(tons))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, species, year)%>%
  dplyr::summarise(ton=sum(tons))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, species)%>%
  dplyr::summarise(ton=mean(ton))
fdi.30$metier=paste(fdi.30$gear_type, fdi.30$target_assemblage, sep='_')

coastal.catch=fdi.30[fdi.30$metier =='OTM_SPF' ,]%>%
  dplyr::group_by(metier, species)%>%
  dplyr::summarise(ton=sum(ton))%>%
  dplyr::arrange(metier, desc(ton))%>%
  dplyr::mutate(cum_val=round(cumsum(ton)/sum(ton), digits=2), prop=ton/sum(ton))

coastal.catch$keep=1
for(i in 2:(nrow(coastal.catch))){
   if(coastal.catch[i,]$metier==coastal.catch[i-1,]$metier){
     if(coastal.catch[i,]$cum_val<=0.99){
       next
     }else{
       coastal.catch[i,]$keep=0
     }
   }else{
     coastal.catch[i,]$keep=1
   }
}
coastal.catch=coastal.catch%>%
  dplyr::filter(keep==1)%>%
  left_join(asfis)

coastal.catch%>%
  ggplot(aes(x=English_name, y=ton))+
  geom_col()+
  facet_wrap(~metier, scales='free_x')+
  xlab('Species name')

#write.csv(coastal.catch, "C:/github/extreme_weather_risk/data/read_only/coastal_catch.csv", row.names = F)
```


