---
title: "Summary of the interviews"
output: pdf_document
date: "2025-04-08"
editor_options: 
  markdown: 
    wrap: 72
---


```{r  , echo=FALSE,message = FALSE}
remove(list=ls())
scriptPath <- rstudioapi::getSourceEditorContext()$path
scriptDir <- dirname(scriptPath)
library(readxl)
library(tidyverse)
library(kableExtra)
coding=read_csv(file.path(scriptDir, '..', 'data/read_only', 'coding_report.csv'))
names(coding)[5]='id_sub'
evts=readxl::read_excel(file.path(scriptDir, '..', 'data', 'values.xlsx'), 
    sheet = "events")
names(evts)[2]='desc'
f.style=coding[coding$id_q==0,c('id_I','value')]
names(f.style)[2]='style'
quest=readxl::read_excel(file.path(scriptDir, '..', 'data', 'values.xlsx'), 
    sheet = "questions")
resp=readxl::read_excel(file.path(scriptDir, '..', 'data', 'values.xlsx'), 
                        sheet = "responses")
#resp$id_I=as.character(resp$id_I)

```

# Results

## Scoring - overview

The questionnaires originated more than 500 scores. About half of the responses (266) were including a clear and unequivocal quantitative score. A large number (123) were not considered relevant by interviewee. Answers that were considered relevant but did not provide a clear score were 150, among which the large majority (105) contained enough information to attempt an interpretation.

```{r , echo=FALSE,message = FALSE,results = 'asis'}
evt.resp=coding[!is.na(coding$id_sub) & !is.na(coding$value),]%>%
  left_join(resp, by=c('id_I', 'id_q','id_sub'))
evt.resp[is.na(evt.resp$score_obs),]$score_obs='900'
evt.resp$fin.score=evt.resp$value
evt.resp$fin.score=ifelse(evt.resp$fin.score==evt.resp$score_obs,evt.resp$fin.score,
                          ifelse(evt.resp$score_obs=='900', ifelse(as.numeric(evt.resp$fin.score)<0,evt.resp$fin.score,'999' ), evt.resp$fin.score))
evt.resp[as.numeric(evt.resp$fin.score)>=0 &as.numeric(evt.resp$fin.score)<=5,]$fin.score='x'

obs=data.frame(table(evt.resp$fin.score))
names(obs)=c('Score','Frequency')

kableExtra::kable(obs%>%arrange(desc(Frequency)), caption='number of answers by interpretation category')%>%
  kable_styling(font_size = 8)%>% 
            column_spec(1, width = "3cm")%>%
            column_spec(2, width = "1.5cm")
```

### Scoring - does not apply (-999)

For each interviewee and event we computed the proportion of the answers scored as -999. These events with more than 50% of answers
scored as -999 were excluded. 
The combination of events/styles excluded are the reported in the table. The image reports the proportion of answers indicating that the event is relevant.

```{r , echo=FALSE,message = FALSE,results = 'asis'}
evt.resp=coding[!is.na(coding$id_sub) & coding$id_q!=21  & !is.na(coding$value),]
evt.resp$type=ifelse(evt.resp$value=='-999' , 'NO' ,'YES')
tab1=evt.resp%>%
  dplyr::group_by(id_I, id_sub, type)%>%
  tally()%>%
  dplyr::group_by(id_I, id_sub)%>%
  dplyr::mutate(prop=n/sum(n))%>%
  #slice_max(prop)%>%
  dplyr::select(id_I, id_sub,type, prop)%>%
  left_join(evts)%>%
  left_join(f.style)
p=ggplot(data = tab1[tab1$type=='YES',])+
  geom_tile(aes(x=desc, y=paste(id_I, style), fill=prop))+
  scale_fill_gradient2(low='red', high = 'blue', midpoint = 0.5)+
  ylab('Interviewee')+
  theme(axis.text.x = element_text(angle=45))
evts.style=evt.resp%>%
  dplyr::group_by(id_I, id_sub, type)%>%
  tally()%>%
  dplyr::group_by(id_I, id_sub)%>%
  dplyr::mutate(prop=n/sum(n))%>%
  slice_max(prop)%>%
  dplyr::filter(type=='NO') # choose which events applies to each style

kableExtra::kable(tab1[tab1$type=='NO' & tab1$prop>0.5, c('desc','style')]%>%replace(is.na(.),0), caption = 'table of the correspondence between fishing styles and relevant extreme weather event')%>%
  kable_styling(font_size = 8)%>% 
            column_spec(1, width = "3cm")%>%
            column_spec(2, width = "1.5cm")
```

```{r , echo=FALSE,message = FALSE,fig.cap='Proportion of responses that were considered relevant by interviewee and event', fig.width=15, fig.height=8}
p
```

\newpage


## Interpretation - Expert scoring

Expert scoring of keywords produced the following results, which were used to assign scores to 87 answers:

```{r , echo=FALSE,message = FALSE,results = 'asis'}

exp.scoring <- read_excel("C:/github/extreme_weather_risk/data/coding_revision_ALL.xlsx")
names(exp.scoring)[6:10]=c('Exp1','Exp2','Exp3','mean','sd')

kable.obj=kableExtra::kable((exp.scoring[,c(3:10)]), booktabs = TRUE, caption = 'scores assigned by experts to answers not providing an explicit scoring') %>%
            kable_styling(font_size = 7)%>% 
            row_spec(0,   bold = TRUE)%>%
            row_spec(1:(nrow(exp.scoring)-1),hline_after=T)%>%
            column_spec(1, width = "2cm")%>%
            column_spec(2, width = "2.5cm")%>%
            column_spec(3, width = "2cm")%>%
            column_spec(4, width = "1.5cm")%>%
   column_spec(5, width = "0.5cm")%>%
   column_spec(6, width = "0.5cm")%>%
   column_spec(7, width = "0.5cm")
        print(kable.obj)
```
\newpage

## Interpretation - AI scoring

This option was applied to a limited number of responses.

**Q1**

*Question*: How much do you think the vary among the group of people you represent? Possible answers: [1–5; from 1 – not at all, to 5 – very
much]

*Observed Answer*: it is so personal, some do not care at all, they just
drink beer all day; we discuss things like this all the time, most are
probably quite stressed by such situations

*Score*: 4

*Why*: The reply indicates clear heterogeneity — a minority who “do not
care at all” vs a majority who are “quite stressed.” That is meaningful
variation but phrased so that most are similar (stressed) with a
minority different. In your training examples, minority exceptions vs
majority agreement tended to map to a high-but-not-maximum value.
Confidence: medium–high.

**Q2**

*Question*: How much do you think the vary among the group of people you
represent? Possible answers: [1–5; from 1 – not at all, to 5 – very
much]

*Observed Answer*: There are some who have rights but don’t practice at
all, and then some who practice like I do, just now and then. Then there
are those who have leased out their permits, and in those cases, there
is somebody who practices quite intensively 

*Score*: 5 

*Why*: The answer describes the full spectrum (no practice → occasional → very intensive).
That explicit broad distribution matches your examples that consistently
received top-end variability scores. Confidence: high. Q2 Question: How
likely is it that you explore new techniques or strategies? Possible
answers: [1–5; from 1 – very unlikely, to 5 – very likely] Observed
Answer: Would probably try to find solutions; activity would certainly
change, and then people would follow that change; there is probably some
variation among people depending on what one is specialized in Score: 4
Why: The speaker says they “would probably try” and expects change and
follow-up by others — language that in your dataset equates to a
positive likelihood but not the absolute “definitely” (which you
typically scored a 5). The note about variation reduces absolute
certainty slightly. Confidence: medium–high. 

**Q3** 

*Question*: If your income
from activity A decreases, can you complement or replace it? Possible
answers: [1–5; from 1 – very unlikely, to 5 – very likely] 

*Observed Answer*: It would be very hard to replace activity A it with other
income. Most have other jobs in winter, but A is such a significant part
of income 

*Score*: 1 

*Why*: The phrase “very hard to replace” and “A is such
a significant part” matches wording in your dataset that mapped to the
low end (very unlikely) — even though there are seasonal other jobs, the
respondent explicitly indicates replacement is essentially not feasible.
Confidence: high.

**Q4** 

*Question*: How easy is it to avoid an extreme weather event by going
to another location?\
Possible answers: [1–5; from 1 – very difficult to avoid, to 5 – very
easy to avoid] 
*Observed Answer*: I Stop my activity Score: 1 Why: The
respondent says they stop the activity instead of relocating to avoid
the event — that implies relocation is not used/feasible. In your
training set, answers that indicate “we don’t/stop” were scored at the
difficult end (1). Confidence: medium — short reply is slightly
ambiguous about whether “stop” means temporarily pause or that
relocation is impossible, but both interpretations point to very low
ease of avoiding by moving. 

**Q5**
*Question*: If the following type of extreme event occurs, how likely is it that you go out and do the
activity? Possible answers: [1–5; from 1 – very unlikely, to 5 – very likely]
*Observed Answer*: we wait because the product can spoil 
*Score*: 1 
*Why*: “We
wait” because of spoilage risk indicates they do not go out during the
event — i.e., very unlikely. In your dataset, similar phrasing (“we wait
/ won’t go because…”) mapped to the lowest likelihood. Confidence: high.

