---
title: "Appendix A2: conditional probability tables."
output:
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r  setup, echo=FALSE,message = FALSE, warning=FALSE}
remove(list=ls())
scriptPath <- rstudioapi::getSourceEditorContext()$path
scriptDir <- dirname(scriptPath)
#setwd(file.path(scriptDir, '..')) 
library(tidyverse)
library(bnlearn)
library(readxl)
library(gRain)
theme_set(theme_bw())
net=read.net(file.path(scriptDir, '..','data/networks/BEWARE_learn_pt2.net'), debug = F)
node.text=read_excel(file.path(scriptDir, '..','data/nodes_text.xlsx'))
#knitr::opts_chunk$set(fig.width=10, fig.height=5) 
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}
head_resp='##'
#target.nodes=c('infrastructure', 'flexibility', 'strategy', 'personal_safety', 'gear', 'target', 'catchability')

```

\setcounter{table}{0}
\renewcommand{\thetable}{B.\arabic{table}}
\renewcommand{\thefigure}{B.\arabic{figure}}

# Network structure

\newpage

# Decision nodes

Controls the climate and fishing management scenarios, those are defined a priori depending on the conditions to be explored in the scenario analyses.

## Extreme weather events (node: event)

The node *Event* is used to wrap all the extreme weather events we are considering.

*Levels (and acronyms)*

-   **Storms** (sto). Strong wind events, with intensity above 24.5 m/s. Season: summer.

-   **Gales** (gal), or moderately strong wind events, with intensity between 15 m/s and 24.5 m/s;

-   **Extreme precipitation** (erf) – rainstorm (summer), or deviation from average precipitation regime;

-   **Algal bloom** (abl): intense algal bloom;

-   **Summer heatwaves** (hws): prolonged discrete anomalously warm water event (Hobday et al., 2016);

-   **Winter heatwaves** (hww): prolonged discrete anomalously warm water event (Hobday et al., 2016);

-   **Icing** (Ici): strong wind events associated to ice, which then forms icing pile up on gears and boats.

## Fishing style (and acronyms)

The node *Fishing style* is used to wrap all the fishing styles and seasons.

*Levels (and acronyms)*

-   **Recreational fisher** (rcf): a fishing style that operates year round. In winter, it primarily performs ice fishing and potentially can fish in open water from the shore; in summer fish with fishing rods both inland and in coastal area;

-   **fishing guide** (fgu): a fishing style that operates year round. In winter, it primarily performs ice fishing and potentially can fish in open water from the shore; in summer fish with fishing rods both inland and in coastal areas;

-   **Coastal small-scale fishers**, (ssf) a summer fishing style that use two kind of gillnets in coastal areas, and in some cases for salmon with fixed traps rigidly regulate. Gillnets types are herring gillnets (gns_spf) and freshwater gillnets (gns_fws). Traps are fpn_ana.

-   **Pelagic trawlers** (trw), represented by pelagic trawlers active in winter and exclusively fishing for small pelagic with midwater trawl (otm);

-   **Archipelago fishers** (arc), which is similar to an artisanal fishers but is by law prevented from selling their catches.

## Fishing management

it controls how much fishing regulations allows fishers to implement adaptation strategies (e.g.: changing target species and gear).

*Levels (and acronyms)*

-   **Business as Usual** (BAU): each fishing style adopts strategies and behaviour according to interviews results

-    **Flexible management** (FLEX). Fishing style is allowed for more flexibility.


## Stock status

it controls how much fishing regulations allows fishers to implement adaptation strategies (e.g.: changing target species and gear).

*Levels (and acronyms)*

-   **status quo** (SQ): each fishing style adopts strategies and behaviour according to interviews results

-    **improved stock status** (FLEX). Fishing style is allowed for more flexibility.

-   **worsened stock status**

\newpage

# Chance nodes - Operational

\newpage
```{r, include=T, echo=FALSE,message=FALSE, warning=FALSE}
target.nodes=unique(node.text[node.text$group=='operational',]$node)
#target.nodes=target.nodes[-grep('risk', target.nodes)]
#target.nodes=c('infrastructure', 'flexibility', 'strategy')
capts.target=c(paste("Figure: Conditional distribution for ", target.nodes))

# plot creation
plot.list=list()
for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # format
  if('fishing_style' %in% names(df.var)){
    df.var=df.var%>%
      dplyr::mutate(fishing_style=ifelse(fishing_style=='recreational',
                                     'rcf', ifelse(fishing_style=='coastal', 'ssf',
                                                   ifelse(fishing_style=='coastal_salmon' , 'sal' ,
                                                           ifelse(fishing_style=='archipelago' ,'arc',
                                                                  ifelse(fishing_style=='trawler' ,'trw','fgu'))))))
  }
  
  # plotting
  if(length(x.parents)==1){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()
    plot.lab='.'
  }
  if(length(x.parents)==2){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'.')
  }
  if(length(x.parents)==3){
   p= ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
        geom_col()+
        facet_grid(cols = vars(!!as.name(x.parents[2])),
               rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
   plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  if(length(x.parents)==4){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])),
                 rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  plot.list[[i]]=p
  names(plot.list)[[i]] = c(paste("conditional distribution for ", target.nodes[i], plot.lab))
}

```


```{r, include=T, echo=FALSE, fig.width = 10, fig.height=5, fig.pos='H', message=FALSE, warning=FALSE, results='asis'}

for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # print characteristics
  cat("\n") 
  cat(sprintf('%s %s\n\n', head_resp, firstup(str_replace(i.node, '_',' '))))
  cat("\n")
  cat(sprintf('_Parents_: %s  \n', paste0(paste(x.parents, collapse = ', '),'.')))
  cat("\n")
  cat(sprintf('_Levels_: %s  \n', paste0(paste(x.levs, collapse = ', '),'.')))
  
  # print description text
  i.text=node.text[node.text$node==i.node,]
  cat("\n")
  cat(sprintf('_Description_: %s  \n', paste0(i.text$text ,'.')))
  
  # plotting
  cat('\n\n') 
  chunk_text <- paste0(
    "```{r, echo=FALSE, fig.width=10, fig.height=5, fig.pos='H', fig.cap='",
    (names(plot.list)[i]),
    "'}\n",
    "print(plot.list[[", i, "]])\n",
    "```\n"
  )
  cat(knitr::knit_child(text = chunk_text, envir = environment(),quiet = TRUE))
  cat("\n\\newpage\n")
  
}
```

\newpage

# Chance nodes - User


```{r, include=T, echo=FALSE,message=FALSE, warning=FALSE}
target.nodes=unique(node.text[node.text$group=='user',]$node)
#target.nodes=target.nodes[-grep('risk', target.nodes)]
#target.nodes=c('infrastructure', 'flexibility', 'strategy')
capts.target=c(paste("Figure: Conditional distribution for ", target.nodes))

# plot creation
plot.list=list()
for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # format
  if('fishing_style' %in% names(df.var)){
    df.var=df.var%>%
      dplyr::mutate(fishing_style=ifelse(fishing_style=='recreational',
                                     'rcf', ifelse(fishing_style=='coastal', 'ssf',
                                                   ifelse(fishing_style=='coastal_salmon' , 'sal' ,
                                                           ifelse(fishing_style=='archipelago' ,'arc',
                                                                  ifelse(fishing_style=='trawler' ,'trw','fgu'))))))
  }
  
  # plotting
  if(length(x.parents)==1){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()
    plot.lab='.'
  }
  if(length(x.parents)==2){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'.')
  }
  if(length(x.parents)==3){
   p= ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
        geom_col()+
        facet_grid(cols = vars(!!as.name(x.parents[2])),
               rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
   plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  if(length(x.parents)==4){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])),
                 rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  plot.list[[i]]=p
  names(plot.list)[[i]] = c(paste("conditional distribution for ", target.nodes[i], plot.lab))
}

```


```{r, include=T, echo=FALSE, fig.width = 10, fig.height=5, fig.pos='H', message=FALSE, warning=FALSE, results='asis'}

for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # print characteristics
  cat("\n") 
  cat(sprintf('%s %s\n\n', head_resp, firstup(str_replace(i.node, '_',' '))))
  cat("\n")
  cat(sprintf('_Parents_: %s  \n', paste0(paste(x.parents, collapse = ', '),'.')))
  cat("\n")
  cat(sprintf('_Levels_: %s  \n', paste0(paste(x.levs, collapse = ', '),'.')))
  
  # print description text
  i.text=node.text[node.text$node==i.node,]
  cat("\n")
  cat(sprintf('_Description_: %s  \n', paste0(i.text$text ,'.')))
  
  # plotting
  cat('\n\n') 
  chunk_text <- paste0(
    "```{r, echo=FALSE, fig.width=10, fig.height=5, fig.pos='H', fig.cap='",
    (names(plot.list)[i]),
    "'}\n",
    "print(plot.list[[", i, "]])\n",
    "```\n"
  )
  cat(knitr::knit_child(text = chunk_text, envir = environment(),quiet = TRUE))
  cat("\n\\newpage\n")
  
}
```


\newpage

# Chance nodes - Economic


```{r, include=T, echo=FALSE,message=FALSE, warning=FALSE}
target.nodes=unique(node.text[node.text$group=='economic',]$node)
#target.nodes=target.nodes[-grep('risk', target.nodes)]
#target.nodes=c('infrastructure', 'flexibility', 'strategy')
capts.target=c(paste("Figure: Conditional distribution for ", target.nodes))

# plot creation
plot.list=list()
for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # format
  if('fishing_style' %in% names(df.var)){
    df.var=df.var%>%
      dplyr::mutate(fishing_style=ifelse(fishing_style=='recreational',
                                     'rcf', ifelse(fishing_style=='coastal', 'ssf',
                                                   ifelse(fishing_style=='coastal_salmon' , 'sal' ,
                                                           ifelse(fishing_style=='archipelago' ,'arc',
                                                                  ifelse(fishing_style=='trawler' ,'trw','fgu'))))))
  }
  
  # plotting
  if(length(x.parents)==1){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()
    plot.lab='.'
  }
  if(length(x.parents)==2){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'.')
  }
  if(length(x.parents)==3){
   p= ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
        geom_col()+
        facet_grid(cols = vars(!!as.name(x.parents[2])),
               rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
   plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  if(length(x.parents)==4){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])),
                 rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  plot.list[[i]]=p
  names(plot.list)[[i]] = c(paste("conditional distribution for ", target.nodes[i], plot.lab))
}

```


```{r, include=T, echo=FALSE, fig.width = 10, fig.height=5, fig.pos='H', message=FALSE, warning=FALSE, results='asis'}

for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # print characteristics
  cat("\n") 
  cat(sprintf('%s %s\n\n', head_resp, firstup(str_replace(i.node, '_',' '))))
  cat("\n")
  cat(sprintf('_Parents_: %s  \n', paste0(paste(x.parents, collapse = ', '),'.')))
  cat("\n")
  cat(sprintf('_Levels_: %s  \n', paste0(paste(x.levs, collapse = ', '),'.')))
  
  # print description text
  i.text=node.text[node.text$node==i.node,]
  cat("\n")
  cat(sprintf('_Description_: %s  \n', paste0(i.text$text ,'.')))
  
  # plotting
  cat('\n\n') 
  chunk_text <- paste0(
    "```{r, echo=FALSE, fig.width=10, fig.height=5, fig.pos='H', fig.cap='",
    (names(plot.list)[i]),
    "'}\n",
    "print(plot.list[[", i, "]])\n",
    "```\n"
  )
  cat(knitr::knit_child(text = chunk_text, envir = environment(),quiet = TRUE))
  cat("\n\\newpage\n")
  
}
```

\newpage

# Chance nodes - Individual


```{r, include=T, echo=FALSE,message=FALSE, warning=FALSE}
target.nodes=unique(node.text[node.text$group=='individual',]$node)
#target.nodes=target.nodes[-grep('risk', target.nodes)]
#target.nodes=c('infrastructure', 'flexibility', 'strategy')
capts.target=c(paste("Figure: Conditional distribution for ", target.nodes))

# plot creation
plot.list=list()
for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # format
  if('fishing_style' %in% names(df.var)){
    df.var=df.var%>%
      dplyr::mutate(fishing_style=ifelse(fishing_style=='recreational',
                                     'rcf', ifelse(fishing_style=='coastal', 'ssf',
                                                   ifelse(fishing_style=='coastal_salmon' , 'sal' ,
                                                           ifelse(fishing_style=='archipelago' ,'arc',
                                                                  ifelse(fishing_style=='trawler' ,'trw','fgu'))))))
  }
  
  # plotting
  if(length(x.parents)==1){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()
    plot.lab='.'
  }
  if(length(x.parents)==2){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'.')
  }
  if(length(x.parents)==3){
   p= ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
        geom_col()+
        facet_grid(cols = vars(!!as.name(x.parents[2])),
               rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
   plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  if(length(x.parents)==4){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])),
                 rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  plot.list[[i]]=p
  names(plot.list)[[i]] = c(paste("conditional distribution for ", target.nodes[i], plot.lab))
}

```


```{r, include=T, echo=FALSE, fig.width = 10, fig.height=5, fig.pos='H', message=FALSE, warning=FALSE, results='asis'}

for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # print characteristics
  cat("\n") 
  cat(sprintf('%s %s\n\n', head_resp, firstup(str_replace(i.node, '_',' '))))
  cat("\n")
  cat(sprintf('_Parents_: %s  \n', paste0(paste(x.parents, collapse = ', '),'.')))
  cat("\n")
  cat(sprintf('_Levels_: %s  \n', paste0(paste(x.levs, collapse = ', '),'.')))
  
  # print description text
  i.text=node.text[node.text$node==i.node,]
  cat("\n")
  cat(sprintf('_Description_: %s  \n', paste0(i.text$text ,'.')))
  
  # plotting
  cat('\n\n') 
  chunk_text <- paste0(
    "```{r, echo=FALSE, fig.width=10, fig.height=5, fig.pos='H', fig.cap='",
    (names(plot.list)[i]),
    "'}\n",
    "print(plot.list[[", i, "]])\n",
    "```\n"
  )
  cat(knitr::knit_child(text = chunk_text, envir = environment(),quiet = TRUE))
  cat("\n\\newpage\n")
  
}
```



\newpage

# Chance nodes -  Societal

```{r, include=T, echo=FALSE,message=FALSE, warning=FALSE}
target.nodes=unique(node.text[node.text$group=='societal',]$node)
#target.nodes=target.nodes[-grep('risk', target.nodes)]
#target.nodes=c('infrastructure', 'flexibility', 'strategy')
capts.target=c(paste("Figure: Conditional distribution for ", target.nodes))

# plot creation
plot.list=list()
for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # format
  if('fishing_style' %in% names(df.var)){
    df.var=df.var%>%
      dplyr::mutate(fishing_style=ifelse(fishing_style=='recreational',
                                     'rcf', ifelse(fishing_style=='coastal', 'ssf',
                                                   ifelse(fishing_style=='coastal_salmon' , 'sal' ,
                                                           ifelse(fishing_style=='archipelago' ,'arc',
                                                                  ifelse(fishing_style=='trawler' ,'trw','fgu'))))))
  }
  
  # plotting
  if(length(x.parents)==1){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()
    plot.lab='.'
  }
  if(length(x.parents)==2){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'.')
  }
  if(length(x.parents)==3){
   p= ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
        geom_col()+
        facet_grid(cols = vars(!!as.name(x.parents[2])),
               rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
   plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  if(length(x.parents)==4){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])),
                 rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  plot.list[[i]]=p
  names(plot.list)[[i]] = c(paste("conditional distribution for ", target.nodes[i], plot.lab))
}

```


```{r, include=T, echo=FALSE, fig.width = 10, fig.height=5, fig.pos='H', message=FALSE, warning=FALSE, results='asis'}

for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # print characteristics
  cat("\n") 
  cat(sprintf('%s %s\n\n', head_resp, firstup(str_replace(i.node, '_',' '))))
  cat("\n")
  cat(sprintf('_Parents_: %s  \n', paste0(paste(x.parents, collapse = ', '),'.')))
  cat("\n")
  cat(sprintf('_Levels_: %s  \n', paste0(paste(x.levs, collapse = ', '),'.')))
  
  # print description text
  i.text=node.text[node.text$node==i.node,]
  cat("\n")
  cat(sprintf('_Description_: %s  \n', paste0(i.text$text ,'.')))
  
  # plotting
  cat('\n\n') 
  chunk_text <- paste0(
    "```{r, echo=FALSE, fig.width=10, fig.height=5, fig.pos='H', fig.cap='",
    (names(plot.list)[i]),
    "'}\n",
    "print(plot.list[[", i, "]])\n",
    "```\n"
  )
  cat(knitr::knit_child(text = chunk_text, envir = environment(),quiet = TRUE))
  cat("\n\\newpage\n")
  
}
```


# Utilities - Risk

Shape how different fisheries will perceive risk. To each of these variables, the effect of the human community variables is to be interpreted as “how much this variable is important”? The CPT table combine the concepts of ranked nodes and the classic definition of Risk = Hazard * Vulnerability. In the case of individual and societal importance, Risk is exactly calculated as Hazard * Vulnerability. Parent categories (Low, medium and high; as well as not importand, somehow important and very important) are mapped to 0, 0.5 and 1 values. The parent value are multiplied.


\newpage
```{r, include=T, echo=FALSE,message=FALSE, warning=FALSE}
target.nodes=unique(node.text[node.text$group=='risk',]$node)
#target.nodes=c('infrastructure', 'flexibility', 'strategy')
capts.target=c(paste("Figure: Conditional distribution for ", target.nodes))

# plot creation
plot.list=list()
for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # format
  if('fishing_style' %in% names(df.var)){
    df.var=df.var%>%
      dplyr::mutate(fishing_style=ifelse(fishing_style=='recreational',
                                     'rcf', ifelse(fishing_style=='coastal', 'ssf',
                                                   ifelse(fishing_style=='coastal_salmon' , 'sal' ,
                                                           ifelse(fishing_style=='archipelago' ,'arc',
                                                                  ifelse(fishing_style=='trawler' ,'trw','fgu'))))))
  }
  
  # plotting
  if(length(x.parents)==1){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()
    plot.lab='.'
  }
  if(length(x.parents)==2){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'.')
  }
  if(length(x.parents)==3){
   p= ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
        geom_col()+
        facet_grid(cols = vars(!!as.name(x.parents[2])),
               rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
   plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  if(length(x.parents)==4){
    p=ggplot(data=df.var, aes_string(x = x.parents[1], y = "Freq", fill = i.node))+
      geom_col()+
      facet_grid(cols = vars(!!as.name(x.parents[2])),
                 rows = vars(!!as.name(x.parents[3])))+
      scale_x_discrete(guide = guide_axis(n.dodge = 2))
    plot.lab=paste0('. Columns represents levels of the node ', x.parents[2],'; rows are levels of the node ', x.parents[3],'.')
  }
  plot.list[[i]]=p
  names(plot.list)[[i]] = c(paste("conditional distribution for ", target.nodes[i], plot.lab))
}

```


```{r, include=T, echo=FALSE, fig.width = 10, fig.height=5, fig.pos='H', message=FALSE, warning=FALSE, results='asis'}

for( i in 1:length(target.nodes)){

  # get values
  i.node=target.nodes[i]
  array.var=net[[i.node]]$prob
  df.var=as.data.frame(array.var)
  xnam=dimnames(array.var)
  x.parents=names(xnam)
  x.parents=x.parents[x.parents!=i.node]
  x.levs=xnam[[i.node]]
  
  # print characteristics
  cat("\n") 
  cat(sprintf('%s %s\n\n', head_resp, firstup(str_replace(i.node, '_',' '))))
  cat("\n")
  cat(sprintf('_Parents_: %s  \n', paste0(paste(x.parents, collapse = ', '),'.')))
  cat("\n")
  cat(sprintf('_Levels_: %s  \n', paste0(paste(x.levs, collapse = ', '),'.')))
  
  # print description text
  i.text=node.text[node.text$node==i.node,]
  cat("\n")
  cat(sprintf('_Description_: %s  \n', paste0(i.text$text ,'.')))
  
  # plotting
  cat('\n\n') 
  chunk_text <- paste0(
    "```{r, echo=FALSE, fig.width=10, fig.height=5, fig.pos='H', fig.cap='",
    (names(plot.list)[i]),
    "'}\n",
    "print(plot.list[[", i, "]])\n",
    "```\n"
  )
  cat(knitr::knit_child(text = chunk_text, envir = environment(),quiet = TRUE))
  cat("\n\\newpage\n")
  
}
```



