---
title: "Appendix A3: literature data treatment."
output:
  pdf_document:
bibliography: references.bib
---

\setcounter{table}{0}
\renewcommand{\thetable}{C.\arabic{table}}
\renewcommand{\thefigure}{C.\arabic{figure}}



# Fishing gears

## Small scale coastal fishers (ssf)

```{r, include=T, echo=F,message=FALSE}
library(tidyverse)
library(readxl)
theme_set(theme_bw())
fdi.eff=read.csv("C:/github/extreme_weather_risk/data/fisheries_statistics/fdi_effort_SWE.csv")
```

We analysed dataset D1, filtered to ICES area 27.3.D.30 and Swedish fleet, to identify the mostly used métiers by the ssf fishers (vessels categories VL0812 and VL0008), where a **métier** is defined as group of fishing activities targeting a similar species or assemblage of species, using similar gear. Figure C.1 shows average total fishing days by métier over the available data (years 2008-2024). We select for consideration in our study the métiers (1) gillnets-freshwater species (GNS-FWS), (2) gillnets-small pelagic species (GNS-SPF), (3) pound nets- anadromous species (FPN-ANA). The sum of average fishing days for the three selected métiers accounts for 85% of total fishing days for ssf in our study area.

```{r, include=T, echo=F,message=FALSE,  results = 'asis', fig.width = 6, fig.height=4, warning=FALSE, fig.cap='Fishing days by metier, swedish small scale coastal fleet in Bothnian Sea'}
ssf.eff=fdi.eff[fdi.eff$sub_region == '27.3.D.30' & fdi.eff$vessel_length_category %in% c('VL0812', 'VL0008'),]


eff.plot=ssf.eff%>%
  dplyr::mutate(total_fishing_days=as.numeric(total_fishing_days))%>%
  dplyr::group_by(fishing_technique,gear_type,target_assemblage,year)%>%
  dplyr::summarise(fdays=sum(total_fishing_days))%>%
  dplyr::group_by(fishing_technique,gear_type, target_assemblage)%>%
  dplyr::summarise(fdays=mean(fdays))

p=eff.plot%>%
  dplyr::filter(fdays>50)%>%
  ggplot(aes(x=target_assemblage, y=fdays/100))+
  geom_col()+
  theme_bw()+
  facet_wrap(~gear_type, scales='free_x')+
  theme(legend.position = 'bottom')+
  xlab('Target species assemblage')+
  ylab('Fishing days (hundreds)')

print(p)
```

```{r, include=T, echo=F,message=FALSE}
## Decision
coastal_fdays=ssf.eff%>%
  dplyr::mutate(total_fishing_days=as.numeric(total_fishing_days))%>%
  dplyr::group_by(fishing_technique,gear_type,target_assemblage,year)%>%
  dplyr::summarise(fdays=sum(total_fishing_days))%>%
  dplyr::group_by(fishing_technique,gear_type, target_assemblage)%>%
  dplyr::summarise(fdays=mean(fdays))%>%
  dplyr::group_by(gear_type, target_assemblage)%>%
  dplyr::summarise(fdays=sum(fdays))%>%
  arrange(desc(fdays))%>%
  ungroup()%>%
  dplyr::mutate(cum_val=cumsum(fdays)/sum(fdays), prop=fdays/sum(fdays)) ## CPT for fishing_style -> gear
coastal_fdays$metier=paste(coastal_fdays$gear_type, coastal_fdays$target_assemblage, sep='_')
write.csv(coastal_fdays, "C:/github/extreme_weather_risk/data/read_only/coastal_fdays.csv", row.names = F)

```

## Pelagic trawlers (trw) and large scale fleet

We analysed dataset D1, filtered to Swedish fleet, to assess the fishing effort distribution by métier in the entire Swedish fleet. This information supports the interpretation of economic data.

```{r, include=T, echo=F,message=FALSE,  results = 'asis', fig.width = 6, fig.height=4, warning=FALSE, fig.cap='Fishing days by metier, swedish large scale fleet'}
lsf.eff=fdi.eff[fdi.eff$vessel_length_category %in% c('VL2440', 'VL40XX'),]


eff.plot=lsf.eff%>%
  dplyr::mutate(total_fishing_days=as.numeric(total_fishing_days))%>%
  dplyr::group_by(fishing_technique,gear_type,target_assemblage,year)%>%
  dplyr::summarise(fdays=sum(total_fishing_days))%>%
  dplyr::group_by(fishing_technique,gear_type, target_assemblage)%>%
  dplyr::summarise(fdays=mean(fdays))

p=eff.plot%>%
  #dplyr::filter(fdays>50)%>%
  ggplot(aes(x=gear_type, y=fdays/100))+
  geom_col()+
  theme_bw()+
  facet_wrap(~fishing_technique, scales='free_x')+
  theme(legend.position = 'bottom')+
  xlab('Target species assemblage')+
  ylab('Fishing days (hundreds)')

print(p)
```


## Recreational fishers

Dataset D3 shows the number of practitioners for each recreational fishing method. Fishing methods based on hook and lines are dominating the distribution reported in Fig. C.3. Cages and nets are the two main techniques tath does not base on hook and line. Cages are predominantly used in crustaceans fishery on the west coast of Sweden. The use of nets in popular in the Baltic Sea. For the purpose of this study we consider that only hook and line methods apply for the recreational fishing style, with no distinction among methods. We apply distinction between inland fishing and sea fishing, using estimates from Table 2 (5814 thousand practitioners inland; 4303 practitioners at sea). Nets applies to the household fishing style.

```{r, include=T, echo=F,message=FALSE,  results = 'asis', fig.width = 6, fig.height=4, warning=FALSE, fig.cap='Fishing days by fishing technique applied by swedish recreational fishers'}
rec.eff=read_excel("C:/github/extreme_weather_risk/data/fisheries_statistics/recf_catch.xlsx", 
                     sheet = "tab6")


eff.plot=rec.eff%>%
  dplyr::group_by(fishing_technique)%>%
  dplyr::summarise(practitioners=sum(practitioners))

p=eff.plot%>%
  #dplyr::filter(fdays>50)%>%
  ggplot(aes(x=fishing_technique, y=practitioners/100))+
  geom_col()+
  theme_bw()+
  #facet_wrap(~gear_type, scales='free_x')+
  theme(legend.position = 'bottom')+
  xlab('Target species assemblage')+
  ylab('Fishing days (hundreds)')

print(p)
```


\newpage


# Catches

## Small scale coastal fishers

We analysed dataset D2 to identify the catch composition associated to ssf métiers identified in the previous section. The dataset, filtered to Swedish fleet in ICES area 27.3.D.30, vessels categories VL0812 and VL0008, was subset to match the ssf métiers (1) gillnets-freshwater species (GNS-FWS), (2) gillnets-small pelagic species (GNS-SPF), (3) pound nets- anadromous species (FPN-ANA). Species scientific name were assigned according to ASFIS List of Species for Fishery Statistics Purposes [@fao2026]. Catch data were aggregated by metier and averaged over years. We calculated the cumulative contribution of each species to a metier and we excluded species falling beyond the 95th percentile. Figure C.3 shows average tons by species and metier over the available data (years 2008-2024). The species mostly represented are Atlantic herring (target of gillnets for small pelagic fish), Atlantic salmon (main target of pound nets for anadromous species), European perch (main target of gillnets for freshwater species). Whitefish is also considered a relevant species as it significantly conbtributes to catches for FPN-ANA and GNS-FWS.


```{r, include=T, echo=F,message=FALSE}
asfis= read_csv("~/ARMELLONI_SLU/miscellaneous_datasets/ASFIS_2020.csv")
names(asfis)[3]='species'
asfis=asfis[,3:5]
fdi.cat=read.csv("C:/github/extreme_weather_risk/data/fisheries_statistics/fdi_catch_SWE.csv")
ssf.catch=fdi.cat[fdi.cat$sub_region == '27.3.D.30' & fdi.cat$vessel_length %in% c('VL0812', 'VL0008'), ]
ssf.eff=read.csv("C:/github/extreme_weather_risk/data/read_only/coastal_fdays.csv")
ssf.catch=ssf.catch%>%
  dplyr::mutate(tons=as.numeric(total_live_weight_landed))%>%
  dplyr::filter(!is.na(tons))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, species, year)%>%
  dplyr::summarise(ton=sum(tons))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, species)%>%
  dplyr::summarise(ton=mean(ton))

```

```{r, include=T, echo=F,message=FALSE,  results = 'asis', fig.width = 6, fig.height=4, warning=FALSE, fig.cap='Catches by metier, swedish small scale coastal fleet in Bothnian Sea (selection of metiers)'}
ssf.catch$metier=paste(ssf.catch$gear_type, ssf.catch$target_assemblage, sep='_')

coastal.catch=ssf.catch[ssf.catch$metier %in% ssf.eff[1:3,]$metier ,]%>%
  dplyr::group_by(metier, species)%>%
  dplyr::summarise(ton=sum(ton))%>%
  dplyr::arrange(metier, desc(ton))%>%
  dplyr::mutate(cum_val=round(cumsum(ton)/sum(ton), digits=2), prop=ton/sum(ton))

coastal.catch$keep=1
for(i in 2:(nrow(coastal.catch))){
   if(coastal.catch[i,]$metier==coastal.catch[i-1,]$metier){
     if(coastal.catch[i,]$cum_val<=0.95){
       next
     }else{
       coastal.catch[i,]$keep=0
     }
   }else{
     coastal.catch[i,]$keep=1
   }
}

coastal.catch=coastal.catch%>%
  dplyr::filter(keep==1)%>%
  left_join(asfis)

p=coastal.catch%>%
  ggplot(aes(y=English_name, x=ton))+
  geom_col()+
  facet_wrap(~metier, scales='free_x')

print(p)
write.csv(coastal.catch, "C:/github/extreme_weather_risk/data/read_only/coastal_catch.csv", row.names = F)

```

## Pelagic trawlers

We analysed dataset D2 to identify the catch composition associated to pelagic trawlers. The dataset, filtered to Swedish fleet in ICES area 27.3.D.30, was subset to match the gear type "OTM". Species scientific name were assigned according to ASFIS List of Species for Fishery Statistics Purposes [@fao2026]. Catch data were aggregated by metier and averaged over years. We calculated the cumulative contribution of each species to a metier and we excluded species falling beyond the 99th percentile. Figure C.4 shows that Atlantic herring contributes for more than 97% of average catches in the study area. For pelagic trawlers we only consider Herring as relevant species.

```{r, include=T, echo=F,message=FALSE}
otm.catch=fdi.cat[fdi.cat$sub_region == '27.3.D.30' & fdi.cat$gear_type %in% c('OTM') & fdi.cat$vessel_length%in% c('VL2440', 'VL40XX'), ]

otm.catch=otm.catch%>%
  dplyr::mutate(tons=as.numeric(total_live_weight_landed))%>%
  dplyr::filter(!is.na(tons))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, species, year)%>%
  dplyr::summarise(ton=sum(tons))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, species)%>%
  dplyr::summarise(ton=mean(ton))

```


```{r, include=T, echo=F,message=FALSE, fig.width = 6, fig.height=4, warning=FALSE, fig.cap='Catches by gear and fishing method, swedish large scale fleet'}
otm.catch.val=fdi.cat[fdi.cat$vessel_length%in% c('VL2440', 'VL40XX'), ]

otm.catch.val=otm.catch.val%>%
  dplyr::mutate(eur=as.numeric(total_value_of_landings))%>%
  dplyr::filter(!is.na(eur))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage, year)%>%
  dplyr::summarise(eur=sum(eur))%>%
  dplyr::group_by(fishing_tech, gear_type, target_assemblage)%>%
  dplyr::summarise(eur=mean(eur))


p=otm.catch.val%>%
  #dplyr::filter(fdays>50)%>%
  ggplot(aes(x=gear_type, y=eur/1000000))+
  geom_col()+
  theme_bw()+
  facet_wrap(~fishing_tech, scales='free_x')+
  theme(legend.position = 'bottom')+
  xlab('Fishing gear')+
  ylab('Fishing value (million €)')
p

```


```{r, , include=T, echo=F,message=FALSE,  results = 'asis', fig.width = 5, fig.height=3, warning=FALSE, fig.cap='Catches for pelagic trawlers, swedish fleet in Bothnian Sea'}
otm.catch$metier=paste(otm.catch$gear_type, otm.catch$target_assemblage, sep='_')

otm.catch=otm.catch%>%
  dplyr::group_by(metier, species)%>%
  dplyr::summarise(ton=sum(ton))%>%
  dplyr::arrange(metier, desc(ton))%>%
  dplyr::mutate(cum_val=round(cumsum(ton)/sum(ton), digits=2), prop=ton/sum(ton))

otm.catch$keep=1
for(i in 2:(nrow(otm.catch))){
   if(otm.catch[i,]$metier==otm.catch[i-1,]$metier){
     if(otm.catch[i,]$cum_val<=0.99){
       next
     }else{
       otm.catch[i,]$keep=0
     }
   }else{
     otm.catch[i,]$keep=1
   }
}

otm.catch=otm.catch%>%
  dplyr::filter(keep==1)%>%
  left_join(asfis)
p=otm.catch%>%
  ggplot(aes(x=English_name, y=ton))+
  geom_col()+
  facet_wrap(~metier, scales='free_x')

print(p)

```

## Recreational fishers

For recreational fishers we are interested in the distribution of number of practitioners by target species.


```{r, include=T, echo=F,message=FALSE,  results = 'asis', fig.width = 6, fig.height=6, warning=FALSE, fig.cap='Catches for recreational fishers'}
rec.catch=read_excel("C:/github/extreme_weather_risk/data/fisheries_statistics/recf_catch.xlsx", 
                     sheet = "tab2")
rec.practics=read_excel("C:/github/extreme_weather_risk/data/fisheries_statistics/recf_catch.xlsx", 
           sheet = "tab1")



rec.catch=rec.catch[rec.catch$name %in% c('Perch', 'Pike', 'Salmon', 'Zander', 'Whitefish', 'Trout', 'Herring','Cod',
                                'Not specified'),]


rec.catch=left_join(rec.catch, rec.practics)
rec.catch$prop=rec.catch$practitioners/sum(rec.catch$practitioners)

rec.catch%>%
  ggplot(aes(y=name, x=practitioners))+
  geom_col()

```
