---
title: "Appendix A3: literature data treatment."
output:
  pdf_document:
    toc: true
    toc_depth: 3
bibliography: references.bib
---


# Economic data

```{r  setup, echo=FALSE,message = FALSE, warning=FALSE}
remove(list=ls())
options(scipen=999)
scriptPath <- rstudioapi::getSourceEditorContext()$path
scriptDir <- dirname(scriptPath)
#setwd(file.path(scriptDir, '..')) 
library(tidyverse)
library(bnlearn)
library(readxl)
library(kableExtra)
library(truncnorm)
theme_set(theme_bw())
int.dat=read_csv(file.path(scriptDir, '..',"data/read_only/coding_report_unc.csv"),  locale = locale(encoding = "ASCII"))
aer.dat=read_csv(file.path(scriptDir, '..',"data/fisheries_statistics/AER_SWE.csv"))
recf.dat=read.csv("C:/github/extreme_weather_risk/data/fisheries_statistics/rec_economics_SWE.csv")
style.dataset=read_excel(file.path(scriptDir, '..',"data/editable_files/dialogues_raw.xlsx"), 
                         sheet = "fishers")

# some formatting
names(style.dataset)[3]='fishing_style'
int.dat=int.dat%>%
  left_join(style.dataset[,c('id_I', 'fishing_style')], by='id_I')


store.cpt=NULL
plot.format=function(xdat){
  
  list.options=str_split(xdat$parent,'\\.')
  xdat$catch=xdat$mitigation=xdat$damage=NA
  
  for(i in 1:length(list.options)){
    opt.vec=list.options[[i]]
    if(length(opt.vec)==2){
      opt.vec=c(opt.vec, 'got')
    }
    xdat[i,c('damage')]=opt.vec[1]
    xdat[i,c('mitigation')]=opt.vec[2]
    xdat[i,c('catch')]=opt.vec[3]
  }
  add.this=data.frame(fishing_style=unique(xdat$fishing_style), parent=NA,
             state='negligible',
             n=1, prob=1, damage='no',
             mitigation='no',
             catch='got')
  xdat=rbind(xdat, add.this)
  return(xdat)
}

```

## Methods

The cost node reflects the economic impact of one single extreme weather event compared to annual average expenses. Cost categories to be used in the conditional probability table are: *negligible* (cost \< 0.05%); *low* (0.05% \< cost \< 2.5%); *medium* (2.5% \< cost \< 5%); *high* (\> 5%). 

The conditional probability tables are estimated through the function `cost.cpt`, here below reported. The input needed are:

 - expenses (mu and sd): annual total expenses, including maintenance, fuel, personnel. This value is anchored in literature, when available;
 
 - fuel.cost (mu and sd): average cost for single fishing trip. This value is anchored in literature, when available;
 
 - damage.small (lo and hi): self declared cost for repairing a minor damage;

 - damage.large (lo and hi): self declared cost for repairing a large damage;
 
 - damage.unc: self declared uncertainty on the cost for repairing damages;
 
 - fuel.avg (lo and hi): self declared fuel consumption for single average fishing trip;
 
 - fuel.large (lo and hi): self declared fuel consumption for single longer-than-average fishing trip;
 
 - miss.catch (mu and sd): average gross economic value of catches from one fishing day
 
 - economic.unc: self declared uncertainty associated with answer regarding expenses;
 
 - fishing style: label describing the fishing style.
 

```{r, include=T, echo=T,results='hide', message=FALSE, warning=FALSE}

cost.cpt=function(expenses.mu, expenses.sd, 
                  fuel.cost.mu,fuel.cost.sd,
         damage.small.lo, damage.small.hi,damage.large.lo, damage.large.hi, 
         damage.unc,
         fuel.avg.lo, fuel.avg.hi, 
         fuel.large.lo, fuel.large.hi,
         miss.catch.mu,miss.catch.sd,
         economic.unc,
         f.style){
  
  # creating distributions 
  ## distribution for annual expenses
  cost.annual=rtruncnorm(a=0, b=Inf, 
                     mean=expenses.mu, 
                     sd=expenses.sd, n=1000) 
  
  ## distribution for minor damage
  cost.lo=rtruncnorm(a=damage.small.lo, b=damage.small.hi, 
                     mean=(damage.small.hi-damage.small.lo)/2, 
                     sd=damage.small.hi*damage.unc, n=1000) 
  
  ## distribution for major damage
  cost.hi=rtruncnorm(a=damage.large.lo, b=damage.large.hi, 
                     mean=(damage.large.hi-damage.large.lo)/2, 
                     sd=damage.large.hi*damage.unc, n=1000) 
  
  ## Distribution for avg travel Unit does not matter as long as it is the same of distribution for large travel.
  avg.fuel=(fuel.avg.hi+fuel.avg.lo)/2
  fuel.lo=rtruncnorm(a=fuel.avg.lo, b=fuel.avg.hi, 
                     mean=avg.fuel, 
                     sd=avg.fuel*economic.unc, n=1000)
  
  ## Distribution for large travel. Unit does not matter as long as it is the same of distribution for avg travel.
  lrg.fuel=(fuel.large.hi+fuel.large.lo)/2
  fuel.hi=rtruncnorm(a=fuel.large.lo, b=fuel.large.hi, 
                     mean=lrg.fuel, 
                     sd=lrg.fuel*economic.unc, n=1000)
  
  # Cost for fuel in extra travel. Multiplies proportion of large to avg travel by monetary cost. 
  extra.fuel.cost=(fuel.hi/fuel.lo)*rtruncnorm(a=0, b=Inf, 
                     mean=fuel.cost.mu, 
                     sd=fuel.cost.sd, n=1000) 
  
  # missing catches
  miss.catch=rtruncnorm(a=0, b=Inf, mean=miss.catch.mu, sd=miss.catch.sd, n=1000)
  
  # create data frame
  cost.df=data.frame(minor.no=cost.lo, major.no=cost.hi, no.travel=extra.fuel.cost)
  cost.df$minor.travel=cost.df$minor.no+cost.df$no.travel
  cost.df$major.travel=cost.df$major.no+cost.df$no.travel
  cost.df.2=cost.df+miss.catch
  names(cost.df.2)=paste(names(cost.df.2), 'miss', sep='.')
  cost.df.2$no.no.miss=miss.catch 
  cost.df=cbind(cost.df, cost.df.2)
  cost.df[,]=apply(cost.df[,],2,function(x) x/cost.annual)
  cost.df.disc=cost.df
  cost.df.disc[,]=apply(cost.df.disc[,],2,
                        function(x) cut(x, c(-Inf, 0.005,0.025,0.05,Inf), 
                                        labels=c('negligible','low','medium','high')))
  
  cost.df.disc$fishing_style=f.style
  cost.cpt=cost.df.disc%>%
  pivot_longer(-fishing_style, names_to = 'parent', values_to = 'state')%>%
  dplyr::group_by(fishing_style,parent, state)%>%
  tally()%>%
  dplyr::mutate(prob=n/sum(n))
  
  return(cost.cpt)
}

error.propagation=function(x.mu, x.er, y.mu, y.er){
  z.mu=x.mu/y.mu
  z.er=sqrt(((x.er/x.mu)^2)+((y.er/y.mu)^2))*z.mu
  return(data.frame(mu=z.mu, sd=z.er))
}

```

\newpage

## Small-scale coastal fishing

Economic information for small scale coastal fishers, available in dataset D3 (Table C.1), filtered for vessel length categories VL0010 and VL0812, fishing technologies FPO and DFN and Swedish country code. Cost of fuel per trip is calculated by dividing total fuel expenses by number of fishing trips. Annual spending is estimated by dividing total annual expenses by number of fishing vessels. Gross value of catches in a single fishing trip is obtained by dividing total annual value of catches by number of fishing trips. Other values are provided in the interviews.

```{r, include=T, echo=F,message=FALSE, results = 'asis', fig.width = 12, fig.height=4, warning=FALSE, fig.cap='CPT estimated for coastal fisherman'}

## economic data from the AER
sek.eur=0.091
x.dat.aer=aer.dat[aer.dat$vessel_length %in% c('VL0010', 'VL0812') 
                & aer.dat$fishing_tech %in% c('FPO','DFN'),]

cost.pt1=x.dat.aer[x.dat.aer$variable_name %in%c('Number of vessels', "Number of fishing trips", 'Energy costs'), 
          c('year', 'vessel_length' ,'fishing_tech', 'variable_name' , 'value')]%>%
  na.omit()%>%
  dplyr::mutate(variable_name=str_remove_all(variable_name, ' '))%>%
  dplyr::filter(year>=2018)%>%
  pivot_wider(names_from = variable_name, values_from = value)%>%
  na.omit()

cost.pt2=x.dat.aer[x.dat.aer$variable_group=='Expenditure',]
cost.pt2=cost.pt2%>%
  dplyr::group_by(vessel_length, fishing_tech,year)%>%
  dplyr::summarise(totaal_exp=sum(value))%>%
  na.omit()%>%
  dplyr::filter(year>=2018)

# cost per unit (trip and vessel)
monetary=left_join(cost.pt1, cost.pt2)
trip.fuel=monetary$Energycosts/monetary$Numberoffishingtrips
annual.spending=monetary$totaal_exp/monetary$Numberofvessels

# value of a missed fishing day
catch.income=x.dat.aer[x.dat.aer$variable_group %in%c('Income'), 
                   c('year', 'vessel_length' ,'fishing_tech', 'variable_name' , 'value')]%>%
  na.omit()%>%
  dplyr::mutate(variable_name=str_remove_all(variable_name, ' '))%>%
  dplyr::filter(year>=2018)%>%
  pivot_wider(names_from = variable_name, values_from = value)%>%
  na.omit()
catch.income$total_income=apply(catch.income[,4:7],1,sum)
catch.income$daily_value=catch.income$Grossvalueoflandings/monetary$Numberoffishingtrips

## economic data from interviews
x.dat=int.dat[int.dat$fishing_style=='small_scale',]
x.dat$range.lwr=as.numeric(x.dat$range.lwr)
x.dat$range.upr=as.numeric(x.dat$range.upr)
damage.unc=mean(x.dat[x.dat$id_q==11,]$uncertainty)
econ.unc=mean(x.dat[x.dat$id_q==39,]$uncertainty)

# cost of damages are already on the same unit we can use them as they are
interview.cost=x.dat[x.dat$short_description%in%c( 'minor_damage_cost', 'major_damage_cost','fuel_trip', 'extra_fuel'),
      c('id_I','short_description', 'range.lwr','range.upr','unit_range')]%>%
  arrange(desc(short_description), id_I)%>%na.omit()

kable.obj=kableExtra::kable((interview.cost), booktabs = TRUE, caption='Variables from interviews') %>%
            kable_styling(font_size = 8)%>% 
            row_spec(0,   bold = TRUE)%>%
            row_spec(1:(nrow(interview.cost)-1),hline_after=T)%>%
            column_spec(1, width = "0.5cm")%>%
            column_spec(2, width = "3cm")%>%
            column_spec(3, width = "1.5cm")%>%
            column_spec(4, width = "1.5cm")%>%
            column_spec(4, width = "2cm")
print(kable.obj)

other.values=data.frame(variable=c('Annual expenses', 'Fuel cost'), 
           value_mean=c(round(mean(annual.spending/sek.eur)/1000), round(mean(trip.fuel/sek.eur))), 
           value_sd=c(round(sd(annual.spending/sek.eur)/1000), round(sd(trip.fuel/sek.eur))), 
           unit=c('thousand SEK','SEK'),
           source=c('STECF_AER', 'STECF_AER'))

kable.obj=kableExtra::kable((other.values), booktabs = TRUE, caption='Variables from literature') %>%
            kable_styling(font_size = 8)%>% 
            row_spec(0,   bold = TRUE)%>%
            row_spec(1:(nrow(other.values)-1),hline_after=T)%>%
            column_spec(1, width = "3cm")%>%
            column_spec(2, width = "1.5cm")%>%
            column_spec(3, width = "1.5cm")%>%
            column_spec(4, width = "1cm")%>%
            column_spec(5, width = "1.5cm")
print(kable.obj)

##
x.cost=cost.cpt(expenses.mu= mean(annual.spending/sek.eur), expenses.sd =sd(annual.spending/sek.eur), 
         fuel.cost.mu = mean(trip.fuel/sek.eur),fuel.cost.sd = sd(trip.fuel/sek.eur),
         damage.small.lo=mean(x.dat[x.dat$short_description=='minor_damage_cost',]$range.lwr), 
         damage.small.hi=mean(x.dat[x.dat$short_description=='minor_damage_cost',]$range.upr),
         damage.large.lo= mean(x.dat[x.dat$short_description=='major_damage_cost',]$range.lwr), 
         damage.large.hi= mean(x.dat[x.dat$short_description=='major_damage_cost',]$range.upr), 
         damage.unc=damage.unc,
         fuel.avg.lo=5, 
         fuel.avg.hi=50, 
         fuel.large.lo=50, 
         fuel.large.hi=100,
         miss.catch.mu = mean(catch.income$daily_value/sek.eur),
         miss.catch.sd = sd(catch.income$daily_value/sek.eur),
         economic.unc=econ.unc,
         f.style = 'small_scale')

plot.format(x.cost)%>%
  ggplot(aes(x=damage, y=prob, fill=state))+
  geom_col()+
  theme(legend.position = 'bottom')+
  facet_grid(cols=vars(mitigation), rows=vars(catch))
store.cpt=rbind(store.cpt, x.cost)
```

\newpage

## Pelagic trawling

Pelagic economic information is derived from dataset D3, filtered for vessel length categories VL2440 and VL40XX, and Sweden country code. Cost of fuel per trip is calculated by dividing total fuel expenses by number of fishing trips. Annual spending is estimated by dividing total annual expenses by number of fishing vessels. The cost of longer travel was not available, and it is based on the assumption that an average travel costs between 80% and 120% of the average fuel cost, while a longer travel is between 100% and 300% of average fuel cost. Other values are provided in the interviews.

```{r, include=T, echo=F, results = 'asis', message=FALSE, fig.width = 12, fig.height=4, warning=FALSE, fig.cap='CPT estimated for pelagic trawlers'}

## economic data from the AER
sek.eur=0.091

x.dat.aer=aer.dat[aer.dat$vessel_length %in% c('VL2440', 'VL40XX'),]

cost.pt1=x.dat.aer[x.dat.aer$variable_name %in%c('Number of vessels', "Number of fishing trips", 'Energy costs'), 
          c('year', 'vessel_length' ,'fishing_tech', 'variable_name' , 'value')]%>%
  na.omit()%>%
  dplyr::mutate(variable_name=str_remove_all(variable_name, ' '))%>%
  dplyr::filter(year>=2018)%>%
  pivot_wider(names_from = variable_name, values_from = value)%>%
  na.omit()

cost.pt2=x.dat.aer[x.dat.aer$variable_group=='Expenditure',]
cost.pt2=cost.pt2%>%
  dplyr::group_by(vessel_length, fishing_tech,year)%>%
  dplyr::summarise(totaal_exp=sum(value))%>%
  na.omit()%>%
  dplyr::filter(year>=2018)

monetary=left_join(cost.pt1, cost.pt2)
trip.fuel=monetary$Energycosts/monetary$Numberoffishingtrips
annual.spending=monetary$totaal_exp/monetary$Numberofvessels

catch.income=x.dat.aer[x.dat.aer$variable_group %in%c('Income'), 
                   c('year', 'vessel_length' ,'fishing_tech', 'variable_name' , 'value')]%>%
  na.omit()%>%
  dplyr::mutate(variable_name=str_remove_all(variable_name, ' '))%>%
  dplyr::filter(year>=2018)%>%
  pivot_wider(names_from = variable_name, values_from = value)%>%
  na.omit()
catch.income$total_income=apply(catch.income[,4:7],1,sum)
catch.income$daily_value=catch.income$Grossvalueoflandings/monetary$Numberoffishingtrips
catch.income$total_income/monetary$Numberofvessels
## economic data from interviews
x.dat=int.dat[int.dat$fishing_style=='trawler',]
x.dat$range.lwr=as.numeric(x.dat$range.lwr)
x.dat$range.upr=as.numeric(x.dat$range.upr)
damage.unc=mean(x.dat[x.dat$id_q==11,]$uncertainty)
econ.unc=mean(x.dat[x.dat$id_q==39,]$uncertainty)

# cost of damages are already on the same unit we can use them as they are
interview.cost=x.dat[x.dat$short_description%in%c( 'minor_damage_cost', 'major_damage_cost','fuel_trip', 'extra_fuel'),
      c('id_I','short_description', 'range.lwr','range.upr','unit_range')]%>%
  arrange(desc(short_description), id_I)

kable.obj=kableExtra::kable((interview.cost), booktabs = TRUE, caption='Variables from interviews') %>%
            kable_styling(font_size = 8)%>% 
            row_spec(0,   bold = TRUE)%>%
            row_spec(1:(nrow(interview.cost)-1),hline_after=T)%>%
            column_spec(1, width = "0.5cm")%>%
            column_spec(2, width = "3cm")%>%
            column_spec(3, width = "1.5cm")%>%
            column_spec(4, width = "1.5cm")%>%
            column_spec(4, width = "2cm")
print(kable.obj)

other.values=data.frame(variable=c('Annual expenses', 'Fuel cost'), 
           value_mean=c(round(mean(annual.spending/sek.eur)/1000), round(mean(trip.fuel/sek.eur))), 
           value_sd=c(round(sd(annual.spending/sek.eur)/1000), round(sd(trip.fuel/sek.eur))), 
           unit=c('thousand SEK','SEK'),
           source=c('STECF_AER', 'STECF_AER'))

kable.obj=kableExtra::kable((other.values), booktabs = TRUE, caption='Variables from literature') %>%
            kable_styling(font_size = 8)%>% 
            row_spec(0,   bold = TRUE)%>%
            row_spec(1:(nrow(other.values)-1),hline_after=T)%>%
            column_spec(1, width = "3cm")%>%
            column_spec(2, width = "1.5cm")%>%
            column_spec(3, width = "1.5cm")%>%
            column_spec(4, width = "1cm")%>%
            column_spec(5, width = "1.5cm")
print(kable.obj)



##
x.cost=cost.cpt(expenses.mu= mean(annual.spending/sek.eur), expenses.sd =sd(annual.spending/sek.eur), 
         fuel.cost.mu = mean(trip.fuel/sek.eur),fuel.cost.sd = sd(trip.fuel/sek.eur),
         damage.small.lo=mean(x.dat[x.dat$short_description=='minor_damage_cost',]$range.lwr), 
         damage.small.hi=mean(x.dat[x.dat$short_description=='minor_damage_cost',]$range.upr),
         damage.large.lo= mean(x.dat[x.dat$short_description=='major_damage_cost',]$range.lwr), 
         damage.large.hi= mean(x.dat[x.dat$short_description=='major_damage_cost',]$range.upr), 
         damage.unc=damage.unc,
         fuel.avg.lo=0.8, 
         fuel.avg.hi=1.2, 
         fuel.large.lo=1, 
         fuel.large.hi=3,
         miss.catch.mu = mean(catch.income$daily_value/sek.eur),
         miss.catch.sd = sd(catch.income$daily_value/sek.eur),
         economic.unc=econ.unc,
         f.style = 'trawler')

plot.format(x.cost)%>%
  ggplot(aes(x=damage, y=prob, fill=state))+
  geom_col()+
  theme(legend.position = 'bottom')+
  facet_grid(cols=vars(mitigation), rows=vars(catch))
store.cpt=rbind(store.cpt, x.cost)
```

\newpage

## Recreational fishing

Pelagic economic information is derived from dataset D6, representing the totality of recreational fisherman in Sweden, with no distinction between region or inland vs coastal fishing. Cost of fuel per trip is separately calculated for travel and for boat fuel. These two costs are then averaged. All the uncertainties (on costs and on number of practitioners) are accounted for by propagating errors. Annual spending is estimated by dividing total annual expenses by number of practitioners. Other values are provided in the interviews.

```{r, include=T, echo=F, results = 'asis',message=FALSE, warning=FALSE, fig.width = 12, fig.height=4, warning=FALSE, fig.cap='CPT estimated for recreational fisherman'}
## economic data from Friskitverkeretet
recf.numb=recf.dat[recf.dat$category=='effort',]
recf.exp=recf.dat[recf.dat$category=='expenses',]

# annual values per person
sek_person=error.propagation(x.mu=recf.exp$mean, x.er=recf.exp$error, 
                             y.mu=recf.numb[1,]$mean/1000, y.er=recf.numb[1,]$error/1000)
sek_person$category=recf.exp$variable

sek_day=error.propagation(x.mu=recf.exp$mean, x.er=recf.exp$error, 
                          y.mu=recf.numb[2,]$mean/1000, y.er=recf.numb[2,]$error/1000)
sek_day$category=recf.exp$variable

sek_boat=error.propagation(x.mu=recf.exp$mean, x.er=recf.exp$error, 
                           y.mu=recf.numb[3,]$mean/1000, y.er=recf.numb[3,]$error/1000)
sek_boat$category=recf.exp$variable

## get mean cost for trips
mean.rf.travel=mean(c(sek_day[sek_day$category %in% c('Expenses for travel'),]$mu,
                      sek_boat[sek_boat$category %in% c('Expenses for boat fuel'),]$mu ))
sd.rf.travel=sqrt(sek_day[sek_day$category %in% c('Expenses for travel'),]$sd^2+sek_boat[sek_boat$category %in% c('Expenses for boat fuel'),]$sd^2)/2


## economic data from interviews
x.dat=int.dat[int.dat$fishing_style=='recreational',]
x.dat$range.lwr=as.numeric(str_remove(x.dat$range.lwr, 'B '))
x.dat$range.upr=as.numeric(str_remove(x.dat$range.upr, 'B '))
damage.unc=mean(x.dat[x.dat$id_q==11,]$uncertainty)
econ.unc=mean(x.dat[x.dat$id_q==39,]$uncertainty)

# cost of damages are already on the same unit we can use them as they are
interview.cost=x.dat[x.dat$short_description%in%c( 'minor_damage_cost', 'major_damage_cost','fuel_trip', 'extra_fuel'),
      c('id_I','short_description', 'range.lwr','range.upr','unit_range')]%>%
  arrange(desc(short_description), id_I)

kable.obj=kableExtra::kable((interview.cost), booktabs = TRUE) %>%
            kable_styling(font_size = 8)%>% 
            row_spec(0,   bold = TRUE)%>%
            row_spec(1:(nrow(interview.cost)-1),hline_after=T)%>%
            column_spec(1, width = "0.5cm")%>%
            column_spec(2, width = "3cm")%>%
            column_spec(3, width = "1.5cm")%>%
            column_spec(4, width = "1.5cm")%>%
            column_spec(4, width = "2cm")
print(kable.obj)

other.values=data.frame(variable=c('Annual expenses', 'Fuel cost'), 
           value_mean=c(round(sek_person[sek_person$category=='Totalt',]$mu), round(mean.rf.travel)), 
           value_sd=c(round(sek_person[sek_person$category=='Totalt',]$sd), round(sd.rf.travel)), 
           unit=c('SEK','SEK'),
           source=c('SWAM', 'SWAM'))

kable.obj=kableExtra::kable((other.values), booktabs = TRUE, caption='Variables from literature') %>%
            kable_styling(font_size = 8)%>% 
            row_spec(0,   bold = TRUE)%>%
            row_spec(1:(nrow(other.values)-1),hline_after=T)%>%
            column_spec(1, width = "3cm")%>%
            column_spec(2, width = "1.5cm")%>%
            column_spec(3, width = "1.5cm")%>%
            column_spec(4, width = "1cm")%>%
            column_spec(5, width = "1.5cm")
print(kable.obj)



##
x.cost=cost.cpt(expenses.mu= sek_person[sek_person$category=='Totalt',]$mu, 
                expenses.sd =sek_person[sek_person$category=='Totalt',]$sd, 
                fuel.cost.mu = mean.rf.travel,fuel.cost.sd = sd.rf.travel,
                damage.small.lo=mean(x.dat[x.dat$short_description=='minor_damage_cost',]$range.lwr), 
                damage.small.hi=mean(x.dat[x.dat$short_description=='minor_damage_cost',]$range.upr),
                damage.large.lo= mean(x.dat[x.dat$short_description=='major_damage_cost',]$range.lwr), 
                damage.large.hi= mean(x.dat[x.dat$short_description=='major_damage_cost',]$range.upr), 
                damage.unc=damage.unc,
                fuel.avg.lo= x.dat[x.dat$short_description=='fuel_trip',]$range.lwr, 
                fuel.avg.hi=x.dat[x.dat$short_description=='fuel_trip',]$range.upr,  
                fuel.large.lo=x.dat[x.dat$short_description=='extra_fuel',]$range.lwr+x.dat[x.dat$short_description=='fuel_trip',]$range.lwr, 
                fuel.large.hi=x.dat[x.dat$short_description=='extra_fuel',]$range.upr+x.dat[x.dat$short_description=='fuel_trip',]$range.upr,
                miss.catch.mu = 0,
                miss.catch.sd = 0, 
                economic.unc=econ.unc,
                f.style = 'recreational')

plot.format(x.cost)%>%
  ggplot(aes(x=damage, y=prob, fill=state))+
  geom_col()+
  theme(legend.position = 'bottom')+
  facet_grid(cols=vars(mitigation), rows=vars(catch))
store.cpt=rbind(store.cpt, x.cost)
```

\newpage

## Fishing guides

No updated public information is available for fishing guide economic aspects in Sweden. All estimations base on interview results and our assumptions. Seasonal expenses are estimated by summing the declared maintenance costs and an estimation of fuel costs.

```{r, include=T, echo=F, results = 'asis',message=FALSE, warning=FALSE, fig.width = 12, fig.height=4, warning=FALSE, fig.cap='CPT estimated for fishing guide'}
## economic data from Friskitverkeretet
## economic data from interviews
x.dat=int.dat[int.dat$fishing_style=='guide',]
x.dat$range.lwr=as.numeric(str_remove(x.dat$range.lwr, 'B '))
x.dat$range.upr=as.numeric(str_remove(x.dat$range.upr, 'B '))
damage.unc=mean(x.dat[x.dat$id_q==11,]$uncertainty)
econ.unc=mean(x.dat[x.dat$id_q==39,]$uncertainty)

# cost of damages are already on the same unit we can use them as they are
interview.cost=x.dat[x.dat$short_description%in%c( 'minor_damage_cost', 'major_damage_cost',
                                                   'fuel_trip', 'extra_fuel', 'season_maint'),
      c('id_I','short_description', 'range.lwr','range.upr','unit_range')]%>%
  arrange(desc(short_description), id_I)

kable.obj=kableExtra::kable((interview.cost), booktabs = TRUE) %>%
            kable_styling(font_size = 8)%>% 
            row_spec(0,   bold = TRUE)%>%
            row_spec(1:(nrow(interview.cost)-1),hline_after=T)%>%
            column_spec(1, width = "0.5cm")%>%
            column_spec(2, width = "3cm")%>%
            column_spec(3, width = "1.5cm")%>%
            column_spec(4, width = "1.5cm")%>%
            column_spec(4, width = "2cm")
print(kable.obj)

# seasonal expenses for this guy is the sum of his own estimation + fuel cost
range.exp.lo=x.dat[x.dat$short_description=='season_maint',]$range.lwr+x.dat[x.dat$short_description=='fdays',]$value*x.dat[x.dat$short_description=='fuel_trip',]$range.lwr
range.exp.hi=(x.dat[x.dat$short_description=='season_maint',]$range.upr*10)+x.dat[x.dat$short_description=='fdays',]$value*x.dat[x.dat$short_description=='fuel_trip',]$range.upr
expenses=rtruncnorm(a=range.exp.lo, b=range.exp.hi, 
           mean=(range.exp.hi-range.exp.lo)/2, 
           sd=range.exp.hi*econ.unc, n=1000) 

other.values=data.frame(variable=c('Annual expenses'), 
           value_mean=c(round(mean(expenses))), 
           value_sd=c(round(sd(expenses))), 
           unit=c('SEK'),
           source=c('Our estimation based on interviews'))

kable.obj=kableExtra::kable((other.values), booktabs = TRUE, caption='Variables from literature') %>%
            kable_styling(font_size = 8)%>% 
            row_spec(0,   bold = TRUE)%>%
            #row_spec(1:(nrow(other.values)-1),hline_after=T)%>%
            column_spec(1, width = "3cm")%>%
            column_spec(2, width = "1.5cm")%>%
            column_spec(3, width = "1.5cm")%>%
            column_spec(4, width = "1cm")%>%
            column_spec(5, width = "1.5cm")
print(kable.obj)


##
x.cost=cost.cpt(expenses.mu= mean(expenses), 
                expenses.sd =sd(expenses), 
                fuel.cost.mu = x.dat[x.dat$short_description=='fuel_trip',]$range.lwr,
                fuel.cost.sd = x.dat[x.dat$short_description=='fuel_trip',]$range.lwr*0.1,
                damage.small.lo=mean(x.dat[x.dat$short_description=='minor_damage_cost',]$range.lwr), 
                damage.small.hi=mean(x.dat[x.dat$short_description=='minor_damage_cost',]$range.upr),
                damage.large.lo= mean(x.dat[x.dat$short_description=='major_damage_cost',]$range.lwr), 
                damage.large.hi= mean(x.dat[x.dat$short_description=='major_damage_cost',]$range.upr), 
                damage.unc=damage.unc,
                fuel.avg.lo= x.dat[x.dat$short_description=='fuel_trip',]$range.lwr*0.8, 
                fuel.avg.hi=x.dat[x.dat$short_description=='fuel_trip',]$range.upr,  
                fuel.large.lo=x.dat[x.dat$short_description=='extra_fuel',]$range.lwr+x.dat[x.dat$short_description=='fuel_trip',]$range.lwr*0.8, 
                fuel.large.hi=x.dat[x.dat$short_description=='extra_fuel',]$range.upr+x.dat[x.dat$short_description=='fuel_trip',]$range.upr, 
                miss.catch.mu = 7000,
                miss.catch.sd = 500, 
                economic.unc=econ.unc,
                f.style = 'guide')

plot.format(x.cost)%>%
  ggplot(aes(x=damage, y=prob, fill=state))+
  geom_col()+
  theme(legend.position = 'bottom')+
  facet_grid(cols=vars(mitigation), rows=vars(catch))
store.cpt=rbind(store.cpt, x.cost)
```

\newpage

## Household fishing

No updated public information is available for archipelago fishers economic aspects in Sweden. All estimations base on interview results and our assumptions. Seasonal expenses are estimated by summing the declared maintenance costs and an estimation of fuel costs.

```{r, include=T, echo=F, results = 'asis',message=FALSE, fig.width = 12, fig.height=4, warning=FALSE, fig.cap='CPT estimated for archipelago fisher'}
x.dat=int.dat[int.dat$fishing_style=='household',]
x.dat$range.lwr=as.numeric(str_remove(x.dat$range.lwr, 'B '))
x.dat$range.upr=as.numeric(str_remove(x.dat$range.upr, 'B '))
damage.unc=mean(x.dat[x.dat$id_q==11,]$uncertainty)
econ.unc=mean(x.dat[x.dat$id_q==39,]$uncertainty)

## forcing this
x.dat[x.dat$short_description=='fdays',]$range.lwr=10
x.dat[x.dat$short_description=='fdays',]$range.upr=20

fuel.lo=100#x.dat[x.dat$short_description=='fuel_trip',]$range.lwr/x.dat[x.dat$short_description=='fdays',]$range.lwr
fuel.hi=300#x.dat[x.dat$short_description=='fuel_trip',]$range.upr/x.dat[x.dat$short_description=='fdays',]$range.lwr


x.dat[x.dat$short_description=='fuel_trip',]$range.lwr=x.dat[x.dat$short_description=='fdays',]$range.lwr*fuel.lo
x.dat[x.dat$short_description=='fuel_trip',]$range.upr=x.dat[x.dat$short_description=='fdays',]$range.upr*fuel.hi


# cost of damages are already on the same unit we can use them as they are
interview.cost=x.dat[x.dat$short_description%in%c( 'minor_damage_cost', 'major_damage_cost',
                                                   'fuel_trip', 'extra_fuel', 'season_maint'),
                     c('id_I','short_description', 'range.lwr','range.upr','unit_range')]%>%
  arrange(desc(short_description), id_I)


kable.obj=kableExtra::kable((interview.cost), booktabs = TRUE) %>%
  kable_styling(font_size = 8)%>% 
  row_spec(0,   bold = TRUE)%>%
  row_spec(1:(nrow(interview.cost)-1),hline_after=T)%>%
  column_spec(1, width = "0.5cm")%>%
  column_spec(2, width = "3cm")%>%
  column_spec(3, width = "1.5cm")%>%
  column_spec(4, width = "1.5cm")%>%
  column_spec(4, width = "2cm")
print(kable.obj)

# seasonal expenses for this guy is the sum of his own estimation + fuel cost




range.exp.lo=x.dat[x.dat$short_description=='season_maint',]$range.upr+x.dat[x.dat$short_description=='fuel_trip',]$range.lwr
range.exp.hi=x.dat[x.dat$short_description=='season_maint',]$range.upr*1.2+x.dat[x.dat$short_description=='fuel_trip',]$range.upr


fuel.mean=(fuel.hi+fuel.lo)/2

expenses=rtruncnorm(a=range.exp.lo, b=range.exp.hi, 
                    mean=(range.exp.hi-range.exp.lo)/2, 
                    sd=range.exp.hi*econ.unc, n=1000) 

other.values=data.frame(variable=c('Annual expenses'), 
           value_mean=c(round(mean(expenses))), 
           value_sd=c(round(sd(expenses))), 
           unit=c('SEK'),
           source=c('Our estimation based on interviews'))

kable.obj=kableExtra::kable((other.values), booktabs = TRUE, caption='Variables from literature') %>%
            kable_styling(font_size = 8)%>% 
            row_spec(0,   bold = TRUE)%>%
            #row_spec(1:(nrow(other.values)-1),hline_after=T)%>%
            column_spec(1, width = "3cm")%>%
            column_spec(2, width = "1.5cm")%>%
            column_spec(3, width = "1.5cm")%>%
            column_spec(4, width = "1cm")%>%
            column_spec(5, width = "1.5cm")
print(kable.obj)


##
x.cost=cost.cpt(expenses.mu= mean(expenses), 
                expenses.sd =sd(expenses), 
                fuel.cost.mu = mean(c(fuel.lo,fuel.hi)),
                fuel.cost.sd = sd(c(fuel.lo,fuel.hi)),
                damage.small.lo=mean(x.dat[x.dat$short_description=='minor_damage_cost',]$range.lwr), 
                damage.small.hi=mean(x.dat[x.dat$short_description=='minor_damage_cost',]$range.upr),
                damage.large.lo= mean(x.dat[x.dat$short_description=='major_damage_cost',]$range.lwr), 
                damage.large.hi= mean(x.dat[x.dat$short_description=='major_damage_cost',]$range.upr), 
                damage.unc=damage.unc,
                fuel.avg.lo= fuel.lo, 
                fuel.avg.hi=fuel.hi, 
                fuel.large.lo=fuel.lo*1.2, 
                fuel.large.hi=fuel.hi*1.2, 
                miss.catch.mu = 0,
                miss.catch.sd = 0, 
                economic.unc=econ.unc,
                f.style = 'household')

plot.format(x.cost)%>%
  ggplot(aes(x=damage, y=prob, fill=state))+
  geom_col()+
  theme(legend.position = 'bottom')+
  facet_grid(cols=vars(mitigation), rows=vars(catch))
store.cpt=rbind(store.cpt, x.cost)


```




```{r, include=T, echo=F, results = 'asis',message=FALSE, warning=FALSE, fig.width = 12, fig.height=4, warning=FALSE, fig.cap='CPT estimated for fishing guide'}

write.csv(store.cpt, file.path(scriptDir, '..',"data/read_only/cost_df_v2.csv"), row.names = F)

```


